---
title: "Tidy Analysis of OSYM Data (WIP)"
author: Alp Biberoglu
date: April 20, 2018
output:
  html_document:
    toc: false
  pdf_document:
    toc: false
    toc_depth: 3
---

```{r,echo=FALSE,results="hide",message=FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

This is an individual assignment, which aims to analyze the ÖSYM data using Tidyverse.



1. Get the highest `max_score` programs from each `exam_type`.

First, we start with setting up the workspace by calling tidyverse and loading the dataframe:

```{r echo=TRUE}
library (tidyverse)
```

```{r echo=TRUE}
load("~/Desktop/ETM 2018 Spring/ETM 58D/Assignments/ÖSYM Individual (20.04)/osym_data_2017_v2.RData")
```

Then, we print the data to have a brief understanding of the format:

```{r echo=TRUE}
print(osym_data_2017)
```

In order to find out max_score programs for each exam type, first we need to find out how many exam types are there using unique() function.

```{r echo=TRUE}
attach(osym_data_2017)
unique(exam_type)
```

Now, let's find the result for exam type MF_4 as an example:

We create a new dataset with just the information of program name, max score and exam type.

```{r echo=TRUE}
osym_data_2017_q1 <- osym_data_2017 %>% select(program_name, max_score, exam_type)
```

Then, we filter this dataset for MF_4

```{r echo=TRUE}
filtered_by_mf4 <- osym_data_2017_q1 %>% filter(exam_type=="MF_4")
```

Now, we arrange this new dataset as per the decreasing order of max_score

```{r echo=TRUE}
filtered_by_mf4 %>% arrange(desc(max_score))
```

Which shows that for MF_4 exam type, the program with highest max_score is Electrical and Electronics Engineering.




