---
title: "Tidy Analysis of OSYM Data"
author: Alp Biberoglu
date: April 20, 2018
output:
  html_document:
    toc: false
  pdf_document:
    toc: false
    toc_depth: 3
---

```{r,echo=FALSE,results="hide",message=FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

This is an individual assignment, which aims to analyze the ÖSYM data using Tidyverse.



1. Get the highest `max_score` programs from each `exam_type`.

First, we start with setting up the workspace by calling tidyverse and loading the dataframe:

```{r}
library (tidyverse)
```

```{r}
load("~/Desktop/ETM 2018 Spring/ETM 58D/Assignments/ÖSYM Individual (20.04)/osym_data_2017_v2.RData")
```

Then, we print the data to have a brief understanding of the format:

```{r}
print(osym_data_2017)
```

In order to find out max_score programs for each exam type, first we need to find out how many exam types are there using unique() function.

```{r}
attach(osym_data_2017)
unique(exam_type)
```

Now, let's find the result for exam type MF_4 as an example:

We create a new dataset with just the information of program name, max score and exam type.

```{r}
osym_data_2017_q1 <- osym_data_2017 %>% select(program_name, max_score, exam_type)
```

Then, we filter this dataset for MF_4

```{r}
filtered_by_mf4 <- osym_data_2017_q1 %>% filter(exam_type=="MF_4")
```

Now, we arrange this new dataset as per the decreasing order of max_score

```{r}
filtered_by_mf4 %>% arrange(desc(max_score))
```

Which shows that for MF_4 exam type, the program with highest max_score is Electrical and Electronics Engineering.

2. Plot the top 10 programs of İSTANBUL ÜNİVERSİTESİ in terms of total quota in a bar chart.

First we filter our original data to obtain all data for just İSTANBUL ÜNİVERSİTESİ, with descending order of "general_quota":

```{r}
osym_data_2017_q2 <- osym_data_2017 %>% filter(university_name == "İSTANBUL ÜNİVERSİTESİ") %>% arrange(desc(as.numeric(general_quota))) %>% slice(1:10)
```

Then, we create the bar graph as follows:

```{r}
ggplot(osym_data_2017_q2, aes(x =program_name, y = as.numeric(general_quota),
fill = general_quota)) + geom_bar(stat = "identity")+ theme( axis.text.x = element_text(angle = 90))
```

3. Calculate the fill rate (sum(`general_placement`)/sum(`general_quota`)) per city and return the top 10 cities.


osym_data_2017_q3 <- osym_data_2017 %>% select(city,general_placement,general_quota) %>% group_by(city) %>% summarise(fill_rate = sum(general_placement)/sum(general_quota)) %>%
  arrange(desc(fill_rate)) %>%
  slice(1:10)


Unfortunately, this code created an error and I couldn't solve it, so I can't show a result for this part.

4. Find full (`general_placement` == `general_quota`) Endüstri Mühendisliği programs (use `grepl`) and draw a scatterplot of min_score vs max_score. Set transparency parameter (`alpha`) to 0.7. Set program colors according to whether it is a foundation university or state university. (Tip: State university programs ids start with 1, foundation 2, KKTC 3 and other abroad 4. You can use `substr` function.).

Now, we should see the difference in using "grepl" rather than regular filtering:

```{r}
ind_eng <- osym_data_2017 %>% filter(grepl("Endüstri Mühendisliği", program_name))
ind_eng_2 <- osym_data_2017 %>% filter(program_name== "Endüstri Mühendisliği")
print(ind_eng)
print(ind_eng_2)
```

With "grepl", we were able to see all programs that include "Endüstri Mühendisliği", however regular filtering just shows the programs which are directly named as "Endüstri Mühendisliği".

After obtaining the data, we find the full programs and make the scatter plot:

```{r}
full_programs <- ind_eng %>% filter(general_quota==general_placement)
print(full_programs)
```

Out of 209 programs, 119 are full.

```{r}
program_id <- as.numeric(program_id)
ggplot(data = full_programs, aes(x = min_score, y = max_score)) +
geom_point(alpha=0.7)
```

Also, among full programs, we can see how many are state and how many are foundation:

```{r}
program_id <- as.numeric(program_id)
devlet <- full_programs %>%  filter(program_id<=100000000)
vakif <- full_programs %>%  filter(program_id<=200000000, program_id>=100000000)
print(devlet)
print(vakif)
```



5. Find the top 10 faculties with the highest quotas and draw a bar chart. Ignore similar names and typos in faculty names.

```{r}
top10 <- osym_data_2017 %>% group_by(faculty_name) %>% summarise(total_quota = sum(as.numeric(general_quota)))
top10 <- top10 %>% arrange(desc(total_quota)) %>% slice(1:10) 
print(top10)
```


6. Find all full medicine programs (Tıp but not Tıp Mühendisliği) of foundation universities group by university calculate total quotas per university and maximum max_score and minimum min_score as bounds, ordered and colored by total quota. (Tip: Use `geom_crossbar`)

We can again use "grepl" to find requested programs:

```{r}
medicals <- osym_data_2017 %>% filter(grepl("Tıp", program_name))  %>% filter(!grepl("Mühendis", program_name))
medicals <- medicals %>% group_by(university_name) %>% summarise(total_quota = sum(as.numeric(general_quota)), max = max(max_score), min = min(min_score))
print(medicals)
```

Now we can plot the graph accordingly:

```{r}
ggplot(medicals) + geom_crossbar(aes(x=university_name, ymin= min, ymax= max, y=500)) + theme( axis.text.x = element_text(angle = 90, size = 5))
```



7. Freestyle: Do an analysis that compares the mechanical engineering (Makine Mühendisliği) and civil engineering (İnşaat Mühendisliği) programs.

```{r}

```

8. Freestyle: Compare Koç University with Bilkent University.

```{r}

```

9. Freestyle: Do your best.

```{r}

```

